#Peer Graded Assignment: Course Project 1

##Analisys

1. Code for reading in the dataset and/or processing the data

```{r}
#obtaining the directory where activity.CSV is
arq <- paste(getwd(),"/Data/activity.CSV",sep="")
#loading activity.CSV into arq
activity <- read.csv(arq, stringsAsFactors = FALSE)
```

2. Histogram of the total number of steps taken each day

```{r}
#aggregating the number of steps by date
steps_by_date <- aggregate(steps~date, activity, FUN = sum)
#generating histogram from number of steps by date
hist(steps_by_date$steps,main = "Total number of steps taken each day",xlab = "steps")
```

3. Mean and median number of steps taken each day

```{r}
#mean
mean(steps_by_date$steps)
#median
median(steps_by_date$steps)
```

4. Time series plot of the average number of steps taken

```{r}
#loding funcionalities from ggplot2
library("ggplot2")
#calculating mean of steps by interval
average_steps <- aggregate(steps~interval, activity, FUN = mean,na.rm=TRUE)
#ploting result through line graph
ggplot (average_steps,aes(interval,steps))+geom_line()+
  xlab("Interval") + ylab("mean steps")
```

5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
#interval with the maximum number of steps
average_steps$interval[average_steps$steps==max(average_steps$steps)]
#number maximum of steps
max(average_steps$steps)
```

6. Code to describe and show a strategy for imputing missing data

```{r}
#number of steps as NA values 
NA_vect <- sum(is.na(activity$steps))
NA_vect

#creating a new dataframe: activity_full
activity_full <- activity
#filling the NA values with the mean of each interval
activity_full$steps[is.na(activity_full$steps)] <- average_steps$steps[activity_full$interval]
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
#aggregating the number of steps by date with no NA values
steps_by_date_no_NA <- aggregate(steps~date, activity_full, FUN = sum)
#histogram generate from dataframe whitout NA values - number of steps by date
hist(steps_by_date_no_NA$steps,main = "Total number of steps taken each day",xlab = "steps")
```

As result of imputing NA values, we can see clearly we have the most part of NA values at the first intervals. From the medium to the ending the values are similar. The analisys take as base the first and second histograms.

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
#converting date type to date type
activity_full$date <- strptime(activity_full$date,"%Y-%m-%d")
#creating a new attibute in activity_full loading a day week name
activity_full$weekdays <- weekdays(activity_full$date)
#attibuiting "weekend" for Sunday and Saturday
activity_full$weekdays [activity_full$weekdays %in% c("Sunday","Saturday")] <- "weekend"
#creating a operator nin that is negation of in
`%nin%` = Negate(`%in%`)
#attibuiting "weekday" for days different from Sunday or Saturday
activity_full$weekdays [activity_full$weekdays %nin% c("weekend")] <- "weekday"
#mean of steps in a table whitout NA values
average_steps_noNA <- aggregate(steps~interval+weekdays, activity_full, FUN = mean,na.rm=TRUE)
#generating plot shared by weekdays
ggplot(average_steps_noNA,aes(interval,steps))+geom_line()+facet_grid(.~weekdays)
```

End of analisys.
